\section{Ideal Solution}

The leader election problem defined in Section~\ref{prob:leader_election} is well studied in
the context of distributed systems. A widely adopted solution is the \textbf{Raft} consensus
algorithm~\cite{4}, which provides strong guarantees of safety and liveness by electing a leader
as part of a replicated state machine protocol.

Raft assumes a partially synchronous system, reliable point-to-point communication channels,
and stable node membership. Leader election in Raft is tightly coupled with log replication and
quorum-based agreement, resulting in well-defined behavior even in the presence of failures.
From a theoretical standpoint, Raft represents an ideal solution for achieving consistent
leadership and consensus in distributed systems.

However, these guarantees come at the cost of significant protocol complexity and communication
overhead. Raft relies on frequent heartbeat exchanges, persistent state, and multiple rounds of
message exchanges per election. Moreover, it assumes the availability of reliable,
connection-oriented communication channels between nodes.

In the context considered in this work—characterized by unreliable broadcast communication,
high message loss, lack of direct addressing, and severely constrained embedded
platforms—the assumptions required by Raft do not hold. As a result, while Raft can be considered
an ideal reference solution in terms of correctness guarantees, it is impractical for the target
environment addressed by this protocol.


% \section{Ideal Solution}
% This problem is not new and there already exists a good solution which is \textbf{RAFT}\cite{4}
% but it has some problem for my use case. First of all it's really an overkill, second it can be
% pretty heavy on the number of message sent.
