\subsubsection{Potential Election}

The \textbf{Potential Election} task is responsible for periodically recomputing the local node
\textbf{potential}[\ref{def:potential}] and determining whether a new leader election should be
initiated.

The task is executed at a frequency derived from a user-defined base value, multiplied by a
node-local \textbf{election time scale factor}[\ref{def:time_scale}]. This factor allows each node
to dynamically adjust the aggressiveness of election attempts based on observed network conditions.

A node initiates a new election when at least one of the following conditions holds:
\begin{enumerate}
  \item the leader heartbeat[\ref{def:heartbeat}] has expired;
  \item the local node potential exceeds the stored leader potential by more than the configured
    tolerance[\ref{def:tolerance}], and no known candidate has a strictly higher potential.
\end{enumerate}

When an election is initiated, the node updates its internal election state as follows:
\begin{itemize}
  \item the number of positive votes is initialized to 1 (self-vote);
  \item the \textbf{election sent} flag is set to true;
  \item the \textbf{voted} flag is set to true;
  \item a \textbf{NewElection}[\ref{message:new_election}] message is broadcast, carrying the node
    identifier, current potential, and current round.
\end{itemize}

If the node initiating the election is \emph{not} the current best candidate[\ref{def:best_candidate}],
the election time scale factor is increased, thereby reducing the frequency of subsequent election
attempts. This mechanism suppresses redundant elections from nodes that are unlikely to win, reducing
network traffic under contention.

In addition, the following state updates are applied:
\begin{itemize}
  \item if the node is currently the leader, both the stored leader and local node potentials are
    updated to the newly computed value;
  \item if the node is the current best candidate[\ref{def:best_candidate}], the stored best
    candidate potential is updated accordingly;
  \item if the best candidate heartbeat[\ref{def:heartbeat}] has expired at the time of election
    initiation, the node adopts itself as the new best candidate and resets the election time scale
    factor to its minimum value.
  \item If neither leader nor best-candidate heartbeats are observed,
    the node resets the election time scale factor to its minimum value.
    This silence-driven reset ensures that election activity resumes aggressively after
    prolonged message loss or leader disappearance, preventing permanent suppression.
\end{itemize}
